Outline paper - Multiple imputation for statistical disclosure control: Creating synthetic datasets with *mice*

### 2021-02-18 mice_synthesizing/paper.Rmd

By default, `mice` imputes all values that are coded as missing in the dataset. However, using the `where` parameter within the `mice`-function, one can choose to impute any value in the dataset. As input, the `where` parameter requires a matrix of the same dimensions as the data, (i.e., a $n \times k$ matrix, with $n$ the number of observations in the collected data) containing logicals $z_{jk}$ that indicate which cells are selected to have their values replaced. This approach allows to *overimpute* a subset of the observed data, or the observed data as a whole, resulting in a dataset that partly or completely consists of synthetic values. Other parameters that are required to be specified are the number of imputations $m$ (`m`), the imputation method (`method`), and the predictor matrix, specifying the predictor variables that are used in the imputation model for every variable $j$ that has some of its values replaced (`predictorMatrix`).

>>>>

By default, `mice` imputes all values that are coded as missing in the dataset. However, using the `where` parameter within the `mice`-function, one can choose to impute any value in the dataset. As input, the `where` parameter requires a matrix of the same dimensions as the data, (i.e., a $n \times k$ matrix, with $n$ the number of observations in the collected data) containing logicals $z_{jk}$ that indicate which cells are selected to have their values replaced. This approach allows to *overimpute* a subset of the observed data, or the observed data as a whole, resulting in a dataset that partly or completely consists of synthetic values. Other parameters that are required to be specified are the imputation method (`method`), the number of imputations ($m$) and the predictor matrix, specifying the predictor variables that are used in the imputation model for every variable $j$ that has some of its values replaced (`predictorMatrix`).

Choosing an adequate method to impute the data is of the utmost importance, as a flawed imputation model will drastically reduce the data utility in further inferences. Therefore, imputation models should be as flexible as possible, as this allows to capture most of the patterns in the data, and to model, possibly unanticipated, data characteristics [@murray_multiple_2018; @rubin_18years_1996]. Parametric methods, albeit easy to implement in practice, may be too restrictive to capture generally complex patterns in the data, especially in the case of nonlinear relations and interactions between multiple variables. Classification and regression trees [CART; @breiman_cart_1984] allow to model more complex patterns in the data, and have therefore been suggested as an appropriate imputation method [@reiter_cart_2005; @burgette_reiter_cart_2010; @doove_buuren_recursive_2014]. Loosely speaking, CART sequentially splits the predictor space into non-overlapping regions in such a way that the within-region variance is as small as possible after every split. As such, CART does not impose any parametric distribution on the data, making it a widely applicable method that allows for a large variety of relationships within the data [@islr_2013]. Given these appealing characteristics and the call for the use of flexible methods when multiply imputing data, we illustrate the performance of `mice` with CART as the specified imputation method.
